---
title: "Data Visuals - 2"
output: 
  html_document:
    toc: yes
    code_folding: show
    theme:
      bg: "#202123"
      fg: "#B8BCC2"
      primary: "#EA80FC"
      base_font:
        google: Ubuntu
      heading_font:
        google: Lato
      version: 3
---

```{r setup, include=FALSE}
if (requireNamespace("thematic")) 
  thematic::thematic_rmd(font = "auto")
```

## Load libraries


```{r, message=FALSE, warning=FALSE}
# install.packages(c("gganimate", "gapminder", "ggridges", "devtools", "png", "gifski", "patchwork"))

library("patchwork") # for making figure panels
library("ggridges")  # for making joyplots 
library("gganimate") # for making animations
library("gapminder") # data available from Gapminder.org 
library("tidyverse") 

diamonds = diamonds
```


# stacked bar chart

```{r, message=FALSE, warning=FALSE}
ggplot(data = diamonds,
       aes(x= cut,
           fill = color)
       )+
  geom_bar()
```




# comparisons means & single data points

```{r, message=FALSE, warning=FALSE}
g = ggplot(
  diamonds,
  aes(x= color,
      y = price)
)

g + 
  stat_summary(fun.data = 'mean_cl_boot',
               color= 'pink',
               size= 1,
               geom = 'pointrange',
               shape= 21,
               fill = 'black'
               )

g +
  geom_point(color= 'orange',
             size= 2,
             alpha= 0.2,
             position = position_jitter(width = 0.1, height = 0)
             )
```



# Violins for bimodal distributions

Violin plots are good when you have many data points and data is continuous. 
```{r, message=FALSE, warning=FALSE}

ggplot(
  diamonds,
  aes(x= color, 
      y = price,
      fill= color)
)+
  geom_violin()
```



# Density ridges

```{r, message=FALSE, warning=FALSE}

ggplot(
  diamonds,
  aes(x= price, 
      y = color,
      fill= color)
)+
  ggridges::geom_density_ridges()

```



# Scatterplot

Scatterplot are good for lookiing at the relationship beyween 2 continuos variables
```{r, message=FALSE, warning=FALSE}

ggplot(
  diamonds,
  aes(x = carat,
      y= price,
      color= color)
)+
  geom_point()
```


# Raster plots

Raster plots are useful for looking how a variable of interest
varies as a function of 2 other variables. 

Example: `carat` values differ from `color` and `clarity`.

```{r, message=FALSE, warning=FALSE}
ggplot(
  diamonds,
  aes(x= color,
      y= clarity,
      z = carat)
)+
  stat_summary2d(fun= 'mean', geom='tile')+
  scale_fill_gradient(low = 'purple' , high = 'yellow')
```



# Temporal data

Line plots are best for temporal data

```{r, message=FALSE, warning=FALSE}

texas = ggplot2::txhousing

texas = texas %>% 
  filter(city %in% c("Dallas", "Fort Worth", "San Antonio", "Houston")) %>% 
  mutate(city = factor(city, levels = c("Dallas", "Houston", "San Antonio", "Fort Worth") ) )

ggplot(
  texas,
  aes(x= year,
      y= median,
      color= city,
      fill= city)
)+
  # stat_summary(fun.data = 'mean_cl_boot',
  #              geom= 'ribbon',
  #              alpha= 0.2,
  #              linetype= 0
  #              )+
  stat_summary(fun='mean', geom = 'line')+
  stat_summary(fun='mean', geom = 'point')
```


















