---
title: "Tidy Modeling in R - 2"
output: 
  html_document:
    toc: yes
    code_folding: show
    theme:
      bg: "#2F2957"
      fg: "#9CFFE7"
      primary: "#68FF1C"
      secondary: "#8D9194"
      base_font:
        google: Lato
      heading_font:
        google: Lato
      # bootswatch: 'minty'
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
if (requireNamespace("thematic")) 
  thematic::thematic_rmd(font = "auto")
```


      
## Tidyverse Principles

Use the magrittr ( %>% ) 

```{r message=FALSE, warning=FALSE}
library(dplyr)
# head(mtcars)

small_mtcars = dplyr::slice( dplyr::arrange(mtcars, gear))
small_mtcars

# more compact way
small_mtcars = slice(arrange(mtcars, gear), 1:10)
small_mtcars
```

## Functional Programming 

Functional programming to iterate over an operation. 

Example: logarithm of the ratio of the fuel efficiency to the car weight

```{r}
# instead of a for loop function

ratios = log(mtcars$mpg / mtcars$wt)
ratios
```

A good approach is to use functions that the output only depends on the inputs and the function has no side effects. Use the `purrr` package for iterative programming.

```{r, message=FALSE, warning=FALSE}
library(purrr)


# the map() function operates on vectors and returns vectors 
# map(vector, function)

# get the square root of mtcars MPG column, showing only 3 rows
map( head(mtcars$mpg, 3), sqrt)
```
Can use anonymous functions inside map functions
```{r}
map2_dbl(mtcars$mpg, mtcars$wt, ~ log(.x / .y)) %>% head()
```


## Tibbles vs Dataframes

Tibbles are more flexible dataframes but also restrictive in the sense that you can't perform some operations which is to prevent serious errors.

```{r}
tbbl = tibble(
  `variable 1` = 1:2, 
  two = 3:4
  )
tbbl


tbbl[ , 2] # or tbbl[ , 'two']
```

# Tidyverse Example 

The dataset is city of Chicago data portal of daily ridership for the city's elevated train stations. This data has 1,088,958 rows and 5 columns.

- station id (numeric)
- station name (char)
- date (char mm/dd/yyyy)
- day of week (char)
- number of riders (numeric)

When making a new column name, format is `new_name` = `old_name`

```{r Example, warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)

# read in the Chicago Train data
# url = "http://bit.ly/raw-train-data-csv"

# write to CSV to save data
# write_csv(all_stations, file = "data/ChicagoTrains.csv")

all_stations = read_csv("data/ChicagoTrains.csv")

glimpse(all_stations)
```

## Select, mutate, group

```{r}

all_stations = all_stations %>% 
  # filter the columns and rename stationname
  dplyr::select(station = stationname, date, rides) %>% 
  # convert dates into date objects
  # convert rides to unit of 1k rides
  dplyr::mutate(date = lubridate::mdy(date), rides = rides/ 1000) %>% 
  # group date and station
  dplyr::group_by(date, station) %>% 
  # summarize the records using the max()
  # .groups = 'drop' drops 2 unused columns station_id and daytype
  summarize(rides = max(rides), .groups = 'drop')
```
























