---
title: "Tidy Modeling in R - 3"
output: 
  html_document:
    toc: yes
    code_folding: show
    theme:
      bg: "#2F2957"
      fg: "#9CFFE7"
      primary: "#68FF1C"
      secondary: "#8D9194"
      base_font:
        google: Lato
      heading_font:
        google: Lato
      # bootswatch: 'minty'
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
if (requireNamespace("thematic")) 
  thematic::thematic_rmd(font = "auto")
```


      
## R Modeling Fundamentals

An example of some fundamentals for modeling in R, using data about crickets (2 species: *O.exclamationis* and *O.niveus*), and the relationship between ambient temperature and the rate of of cricket chirps per minute. Crickets data has 31 rows and 3 columns.

```{r crickets, message=FALSE, warning=FALSE}
library(tidyverse)

# install.packages("modeldata")

# load the data into environment
data(crickets, package = 'modeldata')

# column names of the crickets dataset
names(crickets)

```

### Plot data

Plot the relationship between temperature between species and rate of cricket chirps.

```{r}
library(paletteer)

ggplot(
  data = crickets,
  aes(
    x = temp,
    y = rate,
    color = species,
    pch = species,
    lty = species
  )
)+
  geom_point(size = 2)+
  # simple linear regression model, fit for each species
  geom_smooth(method= lm, formula = NULL, se= FALSE, alpha= 0.5) +
  labs(
    title = "Crickets: rate of chirps & temperature",
    x = "Temp (C)", 
    y = "Chirp Rate (per minute)"
  ) +
  scale_color_paletteer_d(`"awtools::ppalette"`)
```

geom_smooth() arguments:

  - method =`c('lm','glm','gam','loess')`
  - formula = `y ~ x`
  - `se` = display the confidence interval around smooth T/F


The linear regression shows that given temperature the *O.exclamationis* chirps more per minute than the other cricket species. 

A linear regression model (`lm()`) arguments for this function are a) a dataframe and b) model formula.
Formula such as `rate ~ temp` which states the chirp **rate** is the *outcome* and the **temperature** value is the *predictor*. If time were a column in the data then the formula could look like `rate ~ temp + time`, meaning that `temp` and `time` together are the predictors. 


Predictors must be quantitative variables. In order for `rate ~ temp + species` to work, need to use dummy variables. Species in this data has 2 values, can be `0` = "O.exclamationis" and `1` = "O.niveus". In order to prevent issues with the y-intercept and slope for each species, we add an *interaction term*.
To expand all factors to include all possible interactions `rate ~ temp * species`.


Can use in-line functions inside the formula, `rate ~ log(temp)` or `rate ~ I( (temp * 9/5) + 32)` to convert temperature to Celsius. 

```{r}
interaction_fit = lm( rate ~ (temp + species)^2, data = crickets)
interaction_fit
```

**residual** is the vertical distance from the line of best fit a data point is

```{r}
# Have 2 plots next to each other 
par(mfrow = c(1, 2))

# show residuals vs predicted values
plot(interaction_fit, which = 1)
```

Assess if the inclusion of the interaction term in necessary, using the `anova()` method.

```{r}
formula = rate ~ temp + species

main_effect_fit = lm(formula, data = crickets)

# compare the 2
anova(main_effect_fit, interaction_fit)

```
The p-value is 0.25, which implies that there is a lack of evidence against null hypothesis, the interaction term is not needed by the model. 

Residual plots should be re-assessed to make sure that our theoretical assumptions are valid enough to trust the p-values produced by the model.

Use the `summary()` function to inspect the model

```{r}
summary(main_effect_fit)
```

The estimated chirp rate for each species increases by 3.60275 (3.6) chirps as the temperature increases by a single degree. The p-value is statistically significant for this term -10.065 (-10.7), meaning that the species *O.niveus* chirps about 10 fewer chirps per minute compares to the other cricket species. 

If we needed to estimate the chirp rate at a temperature that was not observed in the experiment, we could use the predict() method. 

Let's predict the chirp rate for *O.exclamationis* for temperatures between 15 C and 20 C.

```{r}
new_values = data.frame(
  species = "O. exclamationis",
  temp = 15:20
)

predict(main_effect_fit, newdata = new_values)

```

## Tidy Models format

Tidy Model format style 2 ways of plotting data
```{}
plot(plot_data$x , plot_data$y)

# 1
library(lattice)
xyplot( y ~ x, data= plot_data)

# 2
library(ggplot2)
ggplot(plot_data, aes(x= x, y= y)) + geom_point()
```

If modeling has an outcome with 2 classes, there are a variety of models to use, and to use `predict()`

### modeling functions

| function       | Pkg        | Code                                        |
| :------------- | :--------- | :------------------------------------------ |
| `lda()`        | MASS       | predict(object)                             |
| `glm()`        | stats      | predict(object, type='response')            |
| `gbm()`        | gbm        | predict(object, type = “response”, n.trees) |
| `mda()`        | mda        | predict(object, type = “posterior”)         |
| `rpart()`      | rpart      | predict(object, type = “prob”)              |
| various        | RWeka      | predict(object, type = “probability”)       |
| `logitboost()` | LogitBoost | predict(object, type = “raw”, nIter)        |
| `pamr.train()` | pamr       | pamr.predict(object, type = “posterior”)    |



### Handle missing values

Data that has missing values will cause errors when making predictions
```{r}
# add missing value
new_values$temp[1] = NA

# simply omit the missing value, now 5 rows NOT 6
predict(main_effect_fit, newdata = new_values, na.action = na.omit)
```

### Correlation example

In an example of predictors being used to screen basedon their correlation to the outcome column. 

```{r}
corr_res = map(mtcars %>% select(-mpg), cor.test, y= mtcars$mpg)

corr_res[[1]]


# use the broom library to return a tibble
library(broom)

tidy( corr_res[[1]] )
```


Plot it
```{r}
corr_res %>% 
  # map to a tidy dataframe
  map_dfr(tidy, .id = 'predictor') %>% 
  ggplot( aes(x = fct_reorder(predictor, estimate))) +
  geom_point( aes(y= estimate))+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = .1) +
  labs(x = NULL, y = "Correlation with mpg")
```



### Split data

```{r}
split_by_species = crickets %>% 
  group_nest(species)

# data column has the `rate` and `temp` columns
split_by_species

```

### Model by species

```{r}
# purrr::map()
model_by_species = 
  split_by_species %>% 
  mutate(model = map(data, ~ lm(rate ~ temp, data = .x)))

model_by_species %>% 
  mutate(coef = map(model, tidy)) %>% 
  select(species, coef) %>% 
  unnest(cols = c(coef))
```













