---
title: "Stats & Probability"
output: 
  html_document:
    toc: yes
    code_folding: show
    theme:
      bg: "#202123"
      fg: "#c8cbd0" # grey
      primary: "#00b36b"  # pink "#EA80FC" 
      base_font:
        google: Ubuntu
      heading_font:
        google: Ubuntu
      version: 3
---

Material is lessons from Khan Academy, to refresh my memory and skills.

```{r setup, include=FALSE}
if (requireNamespace("thematic")) 
  thematic::thematic_rmd(font = "auto")
```

```{r, warning=FALSE, message=FALSE ,echo=FALSE}
library(tidyverse)
library(ggvenn)
```



## Frequency Tables & Venn diagrams

```{r}
# library(ggvenn)

circles = list('Bajor' = c(6,3), 'Cardassia' = c(1,3))

ggvenn(circles, set_name_color = 'white', 
       show_elements = T,
       text_color = 'white')
```

- 6 people are Bajoran and not Cardassian 
- 3 people are both Bajoran and Cardassian
- 1 person is Cardassian and not Bajoran


```{r echo=FALSE}
tribble(
  ~' ',~'Bajoran', ~'Cardassian',~'total',
  'Cardassian', 3, 6, 9,
  'Bajoran', 1, 2, 3,
  'total', 4, 8, 12
)

```



## Two way frequency table

| .                | Type A          | Type B           | total |     |
| :--------------- | :-------------- | :--------------- | :---- | :-- |
| Pos X            | 28              | 35               | 60    |     |
| Neg X            | 97              | 104              | 201   |     |
| ---- total ----  | = 125           | = 139            |       |     |
| Proportion Pos X | 28 / 125 = 0.22 | 35 / 139 = 0.25  |       |     |
| Proportion Neg X | 97 / 125 = 0.78 | 104 / 139 = 0.75 |       |     |   


# Descriptive Stats

Average = mean. Median = middle. Mode = most common number

```{r}
a = c(1,2,4,6,4)

sort(a)

total = sum(a)
n = length(a)

# avg = total/n
paste("the avg ", mean(a))

paste("the middle", median(a))

# mode(a)
```



```{r}
x = c(  2,4,12,12 )
y = c( 1,1,2,5,3,2,1,1)

mean(x)
mean(y)


sort(x)

median(x)

sum(x)
length(x)



```

number line 0:20 

dots 
1 2 2 4 3 1 1          ... 1
|-|-|-|-|-|-|-|-|-|-| ...|-|
0 1 2 3 4 5 6 7 8 9 10 ... 19 < distribution >

```{r}
x = c(0,2,4,12,12,5,6,19)
mean(x)

v = c(80,90,92,94,96)
v2 = c(90,92,94,96)

paste(" mean with lowest value: ",mean(v))
paste(" mean without lowest value: ",mean(v2))
```

athlete pulls avg to the right but median will be < mean

3 people make same money 50+50+50/3 =50  median = 50

3 unemployed pull avg left but median will be > mean

x points, remove 1 lowest value , the median move to the right, mean goes up


```{r}
# array has a mean of 4, find missing number
# 

x = c(5,2,2,4,8)

length(x)
mean(x)

r= c(15, 11)
range(r)

x= c(1,0,3,4)
mean(x)
```

```{r}
k = c( )

median(k)
mean(k)
```










